Visualizing demographic evolution using geographically inconsistent census data

### 2019-06-14 paper/study.tex
more text around the results

\subsection{Chicago}
Our first scenario examines Chicago, focusing on a region loosely following the
City's administrative borders. Its demographic composition is well explored in
the literature, with reports of racial divide and
gentrification~\citep{Delmelle2016,Delmelle2017,Hwang2014}. Our goal is
therefore to confirm that our novel method can reproduce existing findings.

The initial state of the prototype is illustrated in Figure~\ref{fig:ui}. The
specific workflow used to identify the existence and details of the
gentrification process is illustrated in Figure~\ref{fig:chiWorkflow}. The first
step is to identify the compositions of each cluster from the boxplots, so
orange is associated with majority of White population, green with majority
Black, and purple with higher proportion of four years of college or more (high
education level). The expanded version of the boxplots for the purple cluster
shows a higher income level and majority of occupations in administrative jobs,
therefore the purple cluster identifies gentrified regions.

The trajectories plot illustrates the process of gentrification, progressively
absorbing regions from the orange cluster (White). This corroborates results
from the literature reporting that Black neighbourhoods are less likely to
gentrify~\citep{Hwang2014}. Moreover, this process is unlikely to be reversed, as
indicated by the limited number of trajectories leaving the purple stream. Next,
we select the region that is gentrified in 2010, by clicking on the
corresponding rectangle in the trajectories plot, updating the information on
the maps and the details portion of the interface.

@@

We consider a region that is approximately the administrative border of the
current city of Toronto, using all seven available aspects with equal weights.
The results are summarised in Figure~\ref{fig:to}, considering eight clusters.

@@

We selected for this scenario a region around the metropolitan area of Los
Angeles (LA), following urban density. The summary of the results using all
aspects with equal weights and eight clusters is illustrated in
Figure~\ref{fig:la}. The full statistical description of the clusters is
illustrated in Figure~\ref{fig:laAll}, where the most relevant aspect of each
cluster is highlighted. From the trajectories plot, we can see that there is a
large but shrinking cluster, depicted in green, one increasing cluster in
purple, an almost constant orange cluster, a smaller but increasing pink
cluster, and three other small clusters. The corresponding map illustrates where
these clusters are located, and that they are somewhat geographically stable,
with some movement between the green, orange, and purple clusters.

@@

>>>>

\subsection{Chicago}
Our first scenario examines Chicago, focusing on a region loosely following the
City's administrative borders. Its demographic composition is well explored in
the literature, with reports of racial divide and
gentrification~\citep{Delmelle2016,Delmelle2017,Hwang2014}, so we expect our
results to contain stable regions where the Race aspect is relevant, and some
degree of population change, with increasing income and education levels.

The initial state of the prototype is illustrated in Figure~\ref{fig:ui}. The
first step is to identify the compositions of each cluster from the boxplots, so
orange is associated with majority of White population, green with majority
Black, and purple with higher proportion of four years of college or more (high
education level). The expanded version of the boxplots for the purple cluster
shows a higher income level and majority of occupations in administrative jobs,
therefore the purple cluster identifies gentrified regions.

The trajectories plot illustrates the process of gentrification, also
illustrated in Figure~\ref{fig:chiWorkflow}, progressively absorbing regions
from the orange cluster (White). This corroborates results from the literature
reporting that Black neighbourhoods are less likely to
gentrify~\citep{Hwang2014}. Moreover, this process appears to be unidirectional,
as indicated by the limited number of trajectories leaving the purple stream.
Next, we select the region that is gentrified in 2010, by clicking on the
corresponding rectangle in the trajectories plot, updating the information on
the maps and the details portion of the interface.

@@

We consider a region that is approximately the administrative border of the
current city of Toronto, using all seven available aspects with equal weights.
While Chicago was fairly stable, Toronto is known to be a more dynamic and
diverse city, with significant and increasing immigrant
population~\citep{hulchanski2007three}. Toronto is also known for a stable and
well defined Jewish community~\citep{Shahar2016}. Therefore, we expect the
combination of stable and dynamic regions on the results, with Place of Birth,
Home Language, and Religion identified as relevant aspects. The results are
summarised in Figure~\ref{fig:to}, considering eight clusters.

@@

We selected for this scenario a region around the metropolitan area of Los
Angeles (LA), following urban density. While Chicago and Toronto demonstrated
the abilities of our method for stable and dynamic cities, LA is considerably
larger, both in population and area. Since our method identifies only the eight
most distinct groups, we expect that some groups not to be identified,
especially if their characterisation is similar to another groups. Further, our
system does not include the census variables related to Hispanic heritage, which
were only included in the newer censuses. However, from the literature, we
expect to see some clusters where Race is an important aspect, including White,
Black, and Asian~\citep{Reibel2003}.

The summary of the results using all aspects with equal weights and eight
clusters is illustrated in Figure~\ref{fig:la}. The full statistical description
of the clusters is illustrated in Figure~\ref{fig:laAll}, where the most
relevant aspect of each cluster is highlighted. From the trajectories plot, we
can see that there is a large but shrinking cluster, depicted in green, one
increasing cluster in purple, an almost constant orange cluster, a smaller but
increasing pink cluster, and three other small clusters. The corresponding map
illustrates where these clusters are located, and that they are somewhat
geographically stable, with some movement between the green, orange, and purple
clusters.

@@

These results are visually similar the analysis presented by
Dellmelle~\citep{Delmelle2016}, what was there identified as 'persistently
struggling' corresponds to the purple cluster in Figure~\ref{fig:la}; some
portions of the 'stable elite' cluster, near Inglewood, correspond to the high
income, high education yellow cluster. The other groups identified by that study
were clumped by our system into the green cluster, which would be further
divided if more than eight clusters were to be used.
